---
title: "Project1 Draft"
author: "Helal El-Zaatari"
date: "September 28, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### We are loading the required packages as well as cleaning the data by subsetting to the relevant columns in our data set (food pounds and food per person). We seperate this subset further by years 2008-2018 in order to do a year by year analysis for every month.




```{r, echo=FALSE}
library(tidyr)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(magrittr)
library(lubridate)
library(ggplot2)

dataset <- read.table("C:\\Users\\Helal\\Desktop\\UMD_Services_Provided_20190719.tsv.txt", fill=TRUE, header=FALSE)

dates = dataset %>% 
  separate(V1, into=c("Month", "Day", "Year")) 
  
fooddata <- dates[,c(1,3,6,7)]
fooddata <- fooddata[-1,]

fooddata$Month <- as.numeric(fooddata$Month)

fooddatabymonth <- fooddata[order(fooddata$Month),]
fooddatabyyear <- fooddata[order(fooddata$Year),]

fooddata2018 <- subset(fooddatabymonth, fooddatabymonth[2]==2018)
fooddata2017 <- subset(fooddatabymonth, fooddatabymonth[2]==2017)
fooddata2016 <- subset(fooddatabymonth, fooddatabymonth[2]==2016)
fooddata2015 <- subset(fooddatabymonth, fooddatabymonth[2]==2015)
fooddata2014 <- subset(fooddatabymonth, fooddatabymonth[2]==2014)
fooddata2013 <- subset(fooddatabymonth, fooddatabymonth[2]==2013)
fooddata2012 <- subset(fooddatabymonth, fooddatabymonth[2]==2012)
fooddata2011 <- subset(fooddatabymonth, fooddatabymonth[2]==2011)
fooddata2010 <- subset(fooddatabymonth, fooddatabymonth[2]==2010)
fooddata2009 <- subset(fooddatabymonth, fooddatabymonth[2]==2009)
fooddata2008 <- subset(fooddatabymonth, fooddatabymonth[2]==2008)
```

### We create tables of the total food pounds per month for every year from 2008 - 2018

```{r, echo=FALSE}
total2018 <- matrix(nrow=12, ncol=2)
total2018[,1] <- c(01,02,03,04,05,06,07,08,09,10,11,12)

for (i in 1:12){
monthsub <- subset(fooddata2018, fooddata2018[,1]==i)
monthsub$V5 <- as.numeric(monthsub$V5)
monthsub <- sum(monthsub[,4])
total2018[i,2] <- monthsub
}

total2018 <- data.frame(total2018)
sum2018 <- sum(total2018[,2])

##### analysis for year 2017 by month ####


total2017 <- matrix(nrow=12, ncol=2)
total2017[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2017, fooddata2017[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2017[i,2] <- monthsub
}

total2017 <- data.frame(total2017)
sum2017 <- sum(total2017[,2])


##### analysis for year 2016 by month ####  

total2016 <- matrix(nrow=12, ncol=2)
total2016[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2016, fooddata2016[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2016[i,2] <- monthsub
}


total2016 <- data.frame(total2016)
sum2016 <- sum(total2016[,2])



##### analysis for year 2015 by month ####

total2015 <- matrix(nrow=12, ncol=2)
total2015[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2015, fooddata2015[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2015[i,2] <- monthsub
}


total2015 <- data.frame(total2015)
sum2015 <- sum(total2015[,2])


##### analysis for year 2014 by month ####

total2014 <- matrix(nrow=12, ncol=2)
total2014[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2014, fooddata2014[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2014[i,2] <- monthsub
}


total2014 <- data.frame(total2014)
sum2014 <- sum(total2014[,2])


##### analysis for year 2013 by month ####

total2013 <- matrix(nrow=12, ncol=2)
total2013[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2013, fooddata2013[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2013[i,2] <- monthsub
}


total2013 <- data.frame(total2013)
sum2013 <- sum(total2013[,2])



##### analysis for year 2012 by month ####

total2012 <- matrix(nrow=12, ncol=2)
total2012[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2012, fooddata2012[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2012[i,2] <- monthsub
}


total2012 <- data.frame(total2012)
sum2012 <- sum(total2012[,2])


##### analysis for year 2011 by month ####

total2011 <- matrix(nrow=12, ncol=2)
total2011[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2011, fooddata2011[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2011[i,2] <- monthsub
}


total2011 <- data.frame(total2011)
sum2011 <- sum(total2011[,2])


##### analysis for year 2010 by month ####

total2010 <- matrix(nrow=12, ncol=2)
total2010[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2010, fooddata2010[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2010[i,2] <- monthsub
}


total2010 <- data.frame(total2010)
sum2010 <- sum(total2010[,2])



##### analysis for year 2009 by month ####

total2009 <- matrix(nrow=12, ncol=2)
total2009[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2009, fooddata2009[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2009[i,2] <- monthsub
}


total2009 <- data.frame(total2009)
sum2009 <- sum(total2009[,2])



##### analysis for year 2008 by month ####

total2008 <- matrix(nrow=12, ncol=2)
total2008[,1] <- c("January","February","March","April","May","June","July","August","September","October","November","December")

for (i in 1:12){
  monthsub <- subset(fooddata2008, fooddata2008[,1]==i)
  monthsub$V5 <- as.numeric(monthsub$V5)
  monthsub <- sum(monthsub[,4])
  total2008[i,2] <- monthsub
}


total2008 <- data.frame(total2008)
sum2008 <- sum(total2008[,2])

```
### we plot the results for a single year 2018 in a bar graph below

```{r, echo=FALSE}
bar2018 <- ggplot(data=total2018, aes(x=  factor(c("January","February","March","April","May","June","July","August","September","October","November","December") , level=c("January","February","March","April","May","June","July","August","September","October","November","December")),y=total2018[,2]) ) +
  geom_bar(stat="identity")

bar2018 + ggtitle("Total Amount of Food per Month from 2018")+
  xlab("Month")+ ylab("Total Amount of Food in Pounds")

```

### We can see from the graph that the month of december has a low amount of food compared to the other months. We want to sum all the food per month for the years 2008-2018 in order to see if we can find any trends for certain months. 

```{r, echol=FALSE}
totaljan <- sum(c(36249,17280,17381,21873,21175,25516,25125,30292,26540,28121,29993))
totalfeb <- sum(c(31687,14618,15745,17801,20389,20216,21428,24264,21895,27419,24890))
totalmarch <- sum(c(31995,13493,24677,23375,24880,20395,20883,24357,28926,31145,29582))
totalapril <- sum(c(31635,15583,21603,22232,21476,19435,24089,29292,30994,28860,25569))
totalmay <-   sum(c(34874,13349,17837,20838,19228,20264,23673,25455,27336,27518,29815))
totaljune <-  sum(c(32474,17287,20854,19727,21088,21479,22520,28988,30601,32337,32512))
totaljuly <- sum(c(32762,19235,18860,21635,19897,17622,26066,29878,31266,28181,27145))
totalaugust <- sum(c(35676,10656,18395,22650,24099,27205,28407,27083,30098,31738,32151))
totalsept <- sum(c(31924,12677,24573,23805,20589,25145,26769,27111,30564,31259,29600))
totaloctober <- sum(c(36292,16403,23287,20192,23279,26226,30634,30151,28980,33213,35334))
totalnov <- sum(c(34368,14544,23911,20971,25453,25306,25263,27657,28553,30504,37027))
totaldec <- sum(c(24267,18652,25628,23208,25625,27623,27021,25713,28310,24589,27552))

totalbymonth <- c(totaljan,totalfeb,totalmarch,totalapril,totalmay,totaljune,totaljuly,totalaugust,totalsept,totaloctober,
                  totalnov,totaldec)
totalbymonth <- data.frame(totalbymonth)

baryear <- ggplot(data=totalbymonth, aes(x=  factor(c("January","February","March","April","May","June","July","August","September","October","November","December") , level=c("January","February","March","April","May","June","July","August","September","October","November","December")),y=totalbymonth)) +
  geom_bar(stat="identity")


baryear + ggtitle("Total Amount of Food per Month from 2008-2018")+
  xlab("Month")+ ylab("Total Amount of Food in Pounds")
```

### we can see from the bar graph that the month of february for some reason always results in less amount of food. Indeed the average per month was 277073 from 2008-2018. THe total amount of food needed for February was 240352 which was approximately 15% lower than the average.
